\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-10-04 papers/methods/paper.tex
Initial rick comments

Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB. For a recent review, see \citet{Huterer2018}. Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.

In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric selected samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae. Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainties will prove to be the limiting factor in our analyses \citep{Betoule2014, Scolnic2017}.

As such, staggering effort is being put into developing more sophisticated supernova cosmology analyses. The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models. First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.  \citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in cosmological parameters as a function of many factors to better remove fitting biases. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either also use simulation corrections or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.

In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance. Section \ref{sec:review} is dedicated to a quick review of the supernova cosmology, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement. We summarise our methodology in Section \ref{sec:conclusion}.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.

This represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population. Accurately characterising this underlying population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology. Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population living in apparent magnitude space. Specifically, for any given supernova our map between absolute magnitude and apparent magnitude may take the traditional form:

@@

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from rearranging equation \ref{eq:standard}, and the cosmologically predicted values. To distinguish them, we denote the observed distance modulus $\mu_{\rm obs}$ and the cosmologically calculated one $\mu_C$. This function is given by

@@

where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review). The cosmological $\mu_C$ is defined as

\begin{align}
\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right],\\
r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
\end{align}
where $r$ is the comoving distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size. Whilst binning data does result in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.

Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data. Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift when supernovae increasingly fall below the detection threshold. An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}. Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data. When using traditional fitting methods, these effects are not built into the likelihood and instead are formed by correcting data, which means their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.

@@

To try and escape the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions. \citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}. In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.

Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects that plague supernova surveys. By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied; we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}. This freedom comes with a cost -- computation. The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion. For ABC methods, we must instead simulate an entire supernova population -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts. This is an intensive process.

@@

Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias. Additionally, tests that have been done have generally been performed on either  $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology and not $w$CDM cosmology. However, quantifying the biases on $w$CDM cosmology with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.

@@

The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by when performing supernova cosmology analyses. In this Section, we outline several of the most prominent challenges.

Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion. Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements. We utilise two dispersion models in this work. The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation in the spectral energy distribution and 30\% from chromatic variation. The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$. These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

The underlying supernova population is further complicated by the presence of outliers. Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae. This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts. Depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers. For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible. However, for photometric surveys, which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit and correct for contaminants.

@@

At the most fundamental level, a supernova cosmology analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function. The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model. In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies. Summaries of observables and model parameters are shown in Table \ref{tab:param_summary} for easy reference.

@@

\subsubsection{Observables}
Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$. Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses. We will also have a probability of each supernovae being a Type Ia, $\hat{p}$. Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus. This is discussed further in Section \ref{sec:systreat}. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate. One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations that supernovae have a non-zero probability of belonging to.

@@

The first layer of the hierarchy represents the parameters that describe each supernova. That is, observed parameters are denoted with a hat, whilst the true (latent) value is denoted without a hat. For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has uncertainty, is different from $c$.

For the moment, let us consider a single supernova and it's classic summary statistics $m_B$, $x_1$, $c$. For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$. A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves and uncertainty. However, this is computational prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$. When using this simplification, our latent variables are given by
\begin{align}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}). \label{eq:pop}
\end{align}
As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including intrinsic dispersion in the supernova model.

@@

With the report summary statistics being biased and their uncertainty under-reported, we face a significant challenge in supernova cosmology. We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model. We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.

>>>>

Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SNIa) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, sako2018sdss}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB. For a recent review, see \citet{Huterer2018}. {\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}

In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric identified samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae. {\green Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as we will required reduced systematic uncertainties to fully utilised increase statistics} \citep{Betoule2014, Scolnic2017}.

As such, staggering effort is being put into developing more sophisticated supernova cosmology analyses. The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models. First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.  \citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of many factors to better remove fitting biases. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.

In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance. Section \ref{sec:review} is dedicated to a quick review of the {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement. We summarise our methodology in Section \ref{sec:conclusion}.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.

This {\green ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$} represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population. Accurately characterising this underlying population, its evolution over redshift, and effects from {\green host-galaxy} environment, is one of the challenges of supernova cosmology. Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population living in apparent magnitude space. Specifically, for any given supernova our map between absolute magnitude and apparent magnitude may take the traditional form:

@@

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from rearranging equation \ref{eq:standard}, and the cosmologically predicted values. We denote the observed distance modulus $\mu_{\rm obs}$ and the cosmologically calculated one $\mu_C$. {\green The $\chi^2$ function minimised is}

@@

where $C^{-1}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Betoule2014} for a review of these uncertainties). The cosmological $\mu_C$ is defined as
{\green 
\begin{align}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right],\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
\end{align}
}
where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size. Whilst binning data results in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.

Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data. Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift when supernovae increasingly fall below the detection threshold. An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}. Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then {\green subtracted from} the observational data. When using traditional fitting methods, these effects are not built into the likelihood and instead are formed by correcting data, which means their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.

@@

To avoid the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions. \citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples. In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.

Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects associated with supernova surveys. By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied; we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}. This freedom comes with a cost -- computation. The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion. For ABC methods, we must instead simulate an entire supernova population -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts. This is an intensive process.

@@

Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias. {\green Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to $w$.}

@@

{\green The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses.} In this Section, we outline several of the most prominent challenges.

{\green Forefront among these challenges is our ignorance of the underlying Type Ia population -- its intrinsic dispersion}. Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements. We utilise two dispersion models in this work. The {\green first is based on the} \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation. The second, {\green based on the} {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. {\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$. These two scatter models are currently {\green assumed} to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

The underlying supernova population is further complicated by the presence of outliers. Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae {\green and represent outliers from the standardised SN Ia sample.} This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts. Depending on the cuts {\green optimising the} ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers. For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible. However, for photometric surveys, which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit and correct for contaminants.

@@

At the most fundamental level, a supernova cosmology analysis is simply a mapping from an underlying population onto an observed population, where {\green cosmological parameters are} encoded directly in the mapping function. The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both {\green the observed and underlying} populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model. In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies. Summaries of observables and model parameters are shown in Table \ref{tab:param_summary} for easy reference.

@@

\subsubsection{Observables}
Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$. Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy {\green stellar }mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses. We will also have a probability of each supernovae being a Type Ia, $\hat{p}$. Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus. This is discussed further in Section \ref{sec:systreat}. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate. One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, {\green these misclassifications} can naturally be modelled and taken into account by introducing secondary populations that supernovae have a non-zero probability of belonging to.

@@

The first layer of the hierarchy represents the parameters that describe each supernova. That is, observed parameters are denoted with a hat, whilst the true (latent) value is denoted without a hat. For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has measurement error, is different from $c$.

For the moment, let us consider a single supernova and it's classic summary statistics $m_B$, $x_1$, $c$. For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$. A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainty. However, this is computational prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$. When using this simplification, our latent variables are given by
\begin{align}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}). \label{eq:pop}
\end{align}
As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including intrinsic dispersion in the supernova model.

@@

With the {\green reported summary statistics $\hat{\eta}$ being biased} and their uncertainty under-reported, we face a significant challenge in supernova cosmology. We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model. We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.
